Randomized trace estimation for parameter-dependent matrices applied to spectral density approximation

### 2025-02-19 paper/spectraldensity.tex
change symbols and write todo list for results

To still apply the bound from \refthm{thm:nystrom-pp} in this scenario, we will have to ensure that the Chebyshev expansion is non-negative. We can think of many ways in which this can be achieved: by using Jackson damping \cite{jackson-1912-approximation-trigonometric,braverman-2022-sublinear-time}, albeit at a significantly worse rate of convergence; by shifting up the expansion $g_{\sigma}^{(m)} + \rho$ with a large enough $\rho > 0$, at the loss of some of the low-rank structure of $g_{\sigma}(t \mtx{I}_n - \mtx{A})$; or by directly minimizing the approximation error over all non-negative Chebyshev polynomials \cite{fejer-1916-uber-trigonometrische}, which realistically only works for small $m$. In the end, we propose the following method to produce a non-negative Chebyshev expansion: Fix an even $m$ and expand the square root of the smoothing kernel in a basis of Chebyshev polynomials up to degree $m/2$
\begin{equation}
    \sqrt{g_{\sigma}}^{(\sfrac{m}{2})}(t - s) = \sum_{l=0}^{m/2} \xi_l(t) T_l(s).
\end{equation}
Then, square this expansion with \refalg{alg:chebyshev-squaring} to obtain the non-negative Chebyshev expansion
\begin{equation}
    \underline{g}^{(m)}_{\sigma}(t - s) = \sum_{l=0}^{m} \mu_l(t) T_l(s).
    \label{equ:non-negative-chebyshev-expansion}
\end{equation}
With a complexity of $\mathcal{O}(m \log(m))$ this procedure is computationally inexpensive and allows for very similar convergence bounds to \reflem{lem:chebyshev-error}, which we will derive in \refsec{subsubsec:chebyshev-nystrom-analysis}. In \reftab{tab:chebyshev-timing-interpolation}, the non-negative expansion can be seen to only be slightly slower than the normal one.

@@

In the end, we propose the following approximation for the smoothed spectral density
\begin{equation}
    \widetilde{\phi}_{\sigma}^{(m)}(t) = \frac{1}{n} \Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}(\underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A}))
    \label{equ:chebyshev-nystrom-formula}
\end{equation}
which is the Nyström++ estimator $\Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}$ \refequ{equ:nystrompp-trace-estimator} applied to the non-negative Chebyshev expansion $\underline{g}_{\sigma}^{(m)}$ \refequ{equ:non-negative-chebyshev-expansion}.

@@

An efficient implementation of the approximation \refequ{equ:chebyshev-nystrom-formula} can be achieved thanks to the invariance of the trace under cyclic permutation of its arguments and the recurrence relation \refequ{equ:chebyshev-recurrence} which the Chebyshev polynomials satisfy. To illustrate this, consider the approximation \refequ{equ:chebyshev-nystrom-formula} when $n_{\mtx{\Psi}}$ is zero, i.e. only a low-rank approximation is computed. Using the definition of the Nyström approximation \refequ{equ:nystrom-approximation}, the cyclic invariance of the trace, and the symmetry of $\mtx{A}$, we can rewrite it as
\begin{equation}
    \frac{1}{n} \Trace((\mtx{\Omega}^{\top} \underline{g}^{(m)}_{\sigma}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega})^{\dagger} (\mtx{\Omega}^{\top} \underline{g}^{(m)}_{\sigma}(t \mtx{I}_n - \mtx{A})^2 \mtx{\Omega})).
    \label{equ:cyclic-property}
\end{equation}
Thanks to the affine form of the Chebyshev expansion, the first parenthesized term in the trace is
\begin{equation}
    \mtx{\Omega}^{\top} \underline{g}^{(m)}_{\sigma}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega} = \sum_{l=0}^{m} \mu_l(t) (\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}),
\end{equation}
meaning we can first precompute the small $n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}$ matrices $\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}$ for $l=0, \dots, m$ using the Chebyshev recurrence \refequ{equ:chebyshev-recurrence} and then inexpensively sum them up with the corresponding coefficients $\mu_l(t)$ for all $t$ in which we want to evaluate the approximation. Thanks to the fast and exact squaring of Chebyshev expansions (cf. \refalg{alg:chebyshev-squaring}), the second parenthesized term in \refequ{equ:cyclic-property} can be approximated in an equivalent way. Unlike expanding the squared smoothing kernel $g_{\sigma}^2$ separately \cite{lin-2017-randomized-estimation}, this approach is consistent, which we have observed to make a noticeable difference in terms of accuracy (cf. \reffig{fig:interpolation-issue}), and which enables provability.

@@

    \Statex \textbf{Output:} Approximate evaluations of the spectral density $\{\widetilde{\phi}_{\sigma}(t_i)\}_{i=1}^{n_t}$
    \State Compute $\{\mu_l(t_i)\}_{l=0}^{m}$ for all $t_i$ with non-negative Chebyshev expansion \refequ{equ:non-negative-chebyshev-expansion}

@@

            \State $\mtx{K}_1(t_i) \gets \mtx{K}_1(t_i) + \mu_l(t_i) \mtx{X}$ \Comment{assemble $\mtx{\Omega}^{\top} g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Omega}$}
            \State $\mtx{L}_1(t_i) \gets \mtx{L}_1(t_i) + \mu_l(t_i) \mtx{Y}$ \Comment{assemble $\mtx{\Omega}^{\top} g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi}$}
            \State $\ell(t_i) \gets \ell(t_i) + \mu_l(t_i) z$ \Comment{assemble $\Trace(\mtx{\Psi}^{\top} g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi})$}

@@

        \State $\mtx{K}_2(t_i) \gets \mtx{K}_2(t_i) + \nu_l(t_i) \mtx{X}$ \Comment{assemble $\mtx{\Omega}^{\top} g_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})^2 \mtx{\Omega}$}

@@

      \State $\widetilde{\phi}_{\sigma}(t_i) \gets \frac{1}{n} \Trace\left( \mtx{K}_1(t_i)^{\dagger}\mtx{K}_2(t_i) \right) + \frac{1}{n n_{\mtx{\Psi}}} \left( \ell(t_i) + \Trace\left( \mtx{L}_1(t_i)^{\top} \mtx{K}_1(t_i)^{\dagger} \mtx{L}_1(t_i) \right)  \right) $ \label{lin:4-nystromchebyshev-nystrom-pp}

@@

What we have found to be crucial for a successful approximation of the full spectral density was the detection of regions wherein lie no eigenvalues, i.e. regions where the spectral density nearly vanishes. For small smoothing parameters $\sigma$, the rapid decay of the Gaussian $g_{\sigma}$ will cause $\mtx{K}_1(t_i) \approx \mtx{\Omega}^{\top} g_{\sigma}(t_i\mtx{I}_n - \mtx{A}) \mtx{\Omega}$ to be close to the zero matrix for parameter values $t_i$ located in such regions. Hence, on \reflin{lin:4-nystromchebyshev-nystrom-pp} we would compute the pseudoinverse of an almost zero matrix, which can be extremely unstable (see \reffig{fig:zerocheck}). Notice that with $\mtx{K}_1(t_i)$ -- a matrix we need to assemble anyway -- we can cheaply compute $\Trace(\mtx{K}_1(t_i)) / n_{\mtx{\Omega}}$, which is the $n_{\mtx{\Omega}}$-query Girard-Hutchinson estimator. We may use it as a rough indicator whether $t_i$ lies within a region where the spectral density nearly vanishes, and subsequently set $\widetilde{\phi}_{\sigma}(t_i) = 0$ instead of computing the full expression on \reflin{lin:4-nystromchebyshev-nystrom-pp}.

@@

\begin{theorem}{Chebyshev-Nyström++ method}{chebyshev-nystrom}
    Let $\phi_{\sigma}$ be the smoothed spectral density of a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ whose spectrum is contained in $[-1, 1]$. Then the Chebyshev-Nyström++ estimate $\widetilde{\phi}_{\sigma}^{(m)}$ with \emph{even} degree $m = \mathcal{O}(\sigma^{-1}(\log(\varepsilon^{-1}) + \log(\sigma^{-1})))$, \emph{even} $n_{\mtx{\Omega}} = n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-1}\log(\delta^{-1})^2)$, and smoothing parameter $0 < \sigma \ll 1$ satisfies with probability at least $1 - \delta$
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \widetilde{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq \varepsilon
        \label{equ:chebyshev-nystrom-error}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for all $\varepsilon \in (0, 1)$ and $\delta \in (0, 1/2)$.
\end{theorem}

\begin{proof}
    The previously derived results can be applied to
    \begin{align}
        &\int_{-1}^{1} \left| \phi_{\sigma}(t) - \widetilde{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \int_{-1}^{1} \left| \phi^{(m)}_{\sigma}(t) - \widetilde{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t && \text{(triangle inequality)} \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \widetilde{\varepsilon} \int_{-1}^{1} \left| \phi^{(m)}_{\sigma}(t) \right|~\mathrm{d}t && \text{(\refthm{thm:nystrom-pp} w.p. $\geq 1 - \delta$)} \notag \\
        &\leq (1 + \widetilde{\varepsilon}) \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \widetilde{\varepsilon} \int_{-1}^{1} \left| \phi_{\sigma}(t) \right|~\mathrm{d}t && \text{(triangle inequality)} \notag \\
        &\leq (1 + \widetilde{\varepsilon}) 2 \underline{E}_{\sigma, m} + \widetilde{\varepsilon}. && \text{(\refthm{thm:chebyshev-error}, normalization)} \notag \\
    \end{align}
    From \reflem{lem:non-negative-chebyshev-error} we see that if $0 < \sigma \ll 1$, choosing $m = \mathcal{O}(\sigma^{-1}(\log(\varepsilon^{-1}) + \log(\sigma^{-1})))$ ensures $\underline{E}_{\sigma, m} \leq \widetilde{\varepsilon}$, and thus, since $\widetilde{\varepsilon} \leq 1$,
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \widetilde{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq (1+\widetilde{\varepsilon}) 2 \widetilde{\varepsilon} + \widetilde{\varepsilon} \leq 5 \widetilde{\varepsilon}
    \end{equation}
    from which the result follows by identifying $\varepsilon = 5 \widetilde{\varepsilon}$.
\end{proof}

>>>>

To still apply the bound from \refthm{thm:nystrom-pp} in this scenario, we will have to ensure that the Chebyshev expansion is non-negative. We can think of many ways in which this can be achieved: by using Jackson damping \cite{jackson-1912-approximation-trigonometric,braverman-2022-sublinear-time}, albeit at a significantly worse rate of convergence; by shifting up the expansion $g_{\sigma}^{(m)} + \rho$ with a large enough $\rho > 0$, at the loss of some of the low-rank structure of $g_{\sigma}(t \mtx{I}_n - \mtx{A})$; or by directly minimizing the approximation error over all non-negative Chebyshev polynomials \cite{fejer-1916-uber-trigonometrische}, which realistically only works for small $m$. In the end, we propose the following method to produce a non-negative Chebyshev expansion: Fix an even $m$ and expand the square root of the smoothing kernel in a basis of Chebyshev polynomials up to degree $m/2$
\begin{equation}
    \sqrt{g_{\sigma}}^{(\sfrac{m}{2})}(t - s) = \sum_{l=0}^{m/2} \xi_l(t) T_l(s).
\end{equation}
Then, square this expansion with \refalg{alg:chebyshev-squaring} to obtain the non-negative Chebyshev expansion
\begin{equation}
    \underline{g}^{(m)}_{\sigma}(t - s) = \sum_{l=0}^{m} \underline{\mu}_l(t) T_l(s).
    \label{equ:non-negative-chebyshev-expansion}
\end{equation}
With a complexity of $\mathcal{O}(m \log(m))$ this procedure is computationally inexpensive and allows for very similar convergence bounds to \reflem{lem:chebyshev-error}, which we will derive in \refsec{subsubsec:chebyshev-nystrom-analysis}. In \reftab{tab:chebyshev-timing-interpolation}, the non-negative expansion can be seen to only be slightly slower than the normal one.

@@

In the end, we propose the following approximation for the smoothed spectral density
\begin{equation}
    \underline{\phi}_{\sigma}^{(m)}(t) = \frac{1}{n} \Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}(\underline{g}_{\sigma}^{(m)}(t \mtx{I}_n - \mtx{A}))
    \label{equ:chebyshev-nystrom-formula}
\end{equation}
which is the Nyström++ estimator $\Nystrpp{\mtx{\Omega}}{\mtx{\Psi}}$ \refequ{equ:nystrompp-trace-estimator} applied to the non-negative Chebyshev expansion $\underline{g}_{\sigma}^{(m)}$ \refequ{equ:non-negative-chebyshev-expansion}.

@@

An efficient implementation of the approximation \refequ{equ:chebyshev-nystrom-formula} can be achieved thanks to the invariance of the trace under cyclic permutation of its arguments and the recurrence relation \refequ{equ:chebyshev-recurrence} which the Chebyshev polynomials satisfy. To illustrate this, consider the approximation \refequ{equ:chebyshev-nystrom-formula} when $n_{\mtx{\Psi}}$ is zero, i.e. only a low-rank approximation is computed. Using the definition of the Nyström approximation \refequ{equ:nystrom-approximation}, the cyclic invariance of the trace, and the symmetry of $\mtx{A}$, we can rewrite it as
\begin{equation}
    \frac{1}{n} \Trace((\mtx{\Omega}^{\top} \underline{g}^{(m)}_{\sigma}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega})^{\dagger} (\mtx{\Omega}^{\top} \underline{g}^{(m)}_{\sigma}(t \mtx{I}_n - \mtx{A})^2 \mtx{\Omega})).
    \label{equ:cyclic-property}
\end{equation}
Thanks to the affine form of the Chebyshev expansion, the first parenthesized term in the trace is
\begin{equation}
    \mtx{\Omega}^{\top} \underline{g}^{(m)}_{\sigma}(t \mtx{I}_n - \mtx{A}) \mtx{\Omega} = \sum_{l=0}^{m} \underline{\mu}_l(t) (\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}),
\end{equation}
meaning we can first precompute the small $n_{\mtx{\Omega}} \times n_{\mtx{\Omega}}$ matrices $\mtx{\Omega}^{\top} T_l(\mtx{A}) \mtx{\Omega}$ for $l=0, \dots, m$ using the Chebyshev recurrence \refequ{equ:chebyshev-recurrence} and then inexpensively sum them up with the corresponding coefficients $\underline{\mu}_l(t)$ for all $t$ in which we want to evaluate the approximation. Thanks to the fast and exact squaring of Chebyshev expansions (cf. \refalg{alg:chebyshev-squaring}), the second parenthesized term in \refequ{equ:cyclic-property} can be approximated in an equivalent way. Unlike expanding the squared smoothing kernel $g_{\sigma}^2$ separately \cite{lin-2017-randomized-estimation}, this approach is consistent, which we have observed to make a noticeable difference in terms of accuracy (cf. \reffig{fig:interpolation-issue}), and which enables provability.

@@

    \Statex \textbf{Output:} Approximate evaluations of the spectral density $\{\underline{\phi}_{\sigma}^{(m)}(t_i)\}_{i=1}^{n_t}$
    \State Compute $\{\underline{\mu}_l(t_i)\}_{l=0}^{m}$ for all $t_i$ with non-negative Chebyshev expansion \refequ{equ:non-negative-chebyshev-expansion}

@@

            \State $\mtx{K}_1(t_i) \gets \mtx{K}_1(t_i) + \underline{\mu}_l(t_i) \mtx{X}$ \Comment{assemble $\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Omega}$}
            \State $\mtx{L}_1(t_i) \gets \mtx{L}_1(t_i) + \underline{\mu}_l(t_i) \mtx{Y}$ \Comment{assemble $\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi}$}
            \State $\ell(t_i) \gets \ell(t_i) + \underline{\mu}_l(t_i) z$ \Comment{assemble $\Trace(\mtx{\Psi}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A}) \mtx{\Psi})$}

@@

        \State $\mtx{K}_2(t_i) \gets \mtx{K}_2(t_i) + \nu_l(t_i) \mtx{X}$ \Comment{assemble $\mtx{\Omega}^{\top} \underline{g}_{\sigma}^{(m)}(t\mtx{I}_n - \mtx{A})^2 \mtx{\Omega}$}

@@

      \State $\underline{\phi}_{\sigma}(t_i) \gets \frac{1}{n} \Trace\left( \mtx{K}_1(t_i)^{\dagger}\mtx{K}_2(t_i) \right) + \frac{1}{n n_{\mtx{\Psi}}} \left( \ell(t_i) + \Trace\left( \mtx{L}_1(t_i)^{\top} \mtx{K}_1(t_i)^{\dagger} \mtx{L}_1(t_i) \right)  \right) $ \label{lin:4-nystromchebyshev-nystrom-pp}

@@

What we have found to be crucial for a successful approximation of the full spectral density was the detection of regions wherein lie no eigenvalues, i.e. regions where the spectral density nearly vanishes. For small smoothing parameters $\sigma$, the rapid decay of the Gaussian $g_{\sigma}$ will cause $\mtx{K}_1(t_i) \approx \mtx{\Omega}^{\top} g_{\sigma}(t_i\mtx{I}_n - \mtx{A}) \mtx{\Omega}$ to be close to the zero matrix for parameter values $t_i$ located in such regions. Hence, on \reflin{lin:4-nystromchebyshev-nystrom-pp} we would compute the pseudoinverse of an almost zero matrix, which can be extremely unstable (see \reffig{fig:zerocheck}). Notice that with $\mtx{K}_1(t_i)$ -- a matrix we need to assemble anyway -- we can cheaply compute $\Trace(\mtx{K}_1(t_i)) / n_{\mtx{\Omega}}$, which is the $n_{\mtx{\Omega}}$-query Girard-Hutchinson estimator. We may use it as a rough indicator whether $t_i$ lies within a region where the spectral density nearly vanishes, and subsequently set $\underline{\phi}^{(m)}_{\sigma}(t_i) = 0$ instead of computing the full expression on \reflin{lin:4-nystromchebyshev-nystrom-pp}.

@@

\begin{theorem}{Chebyshev-Nyström++ method}{chebyshev-nystrom}
    Let $\phi_{\sigma}$ be the smoothed spectral density of a symmetric matrix $\mtx{A} \in \mathbb{R}^{n \times n}$ whose spectrum is contained in $[-1, 1]$. Then the Chebyshev-Nyström++ estimate $\underline{\phi}_{\sigma}^{(m)}$ with \emph{even} degree $m = \mathcal{O}(\sigma^{-1}(\log(\varepsilon^{-1}) + \log(\sigma^{-1})))$, \emph{even} $n_{\mtx{\Omega}} = n_{\mtx{\Psi}} = \mathcal{O}(\varepsilon^{-1}\log(\delta^{-1})^2)$, and smoothing parameter $0 < \sigma \ll 1$ satisfies with probability at least $1 - \delta$
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq \varepsilon
        \label{equ:chebyshev-nystrom-error}
        %\lVert  \phi_{\sigma} - \phi_{\sigma}^{(m)} \rVert _1 &\leq \frac{2\sqrt{2}}{\sigma^2} (1 + \sigma)^{-m}.
        %\label{equ:2-chebyshev-interpolation-error}
    \end{equation}
    for all $\varepsilon \in (0, 1)$ and $\delta \in (0, 1/2)$.
\end{theorem}

\begin{proof}
    The previously derived results can be applied to
    \begin{align}
        &\int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \int_{-1}^{1} \left| \phi^{(m)}_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t && \text{(triangle inequality)} \notag \\
        &\leq \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \widetilde{\varepsilon} \int_{-1}^{1} \left| \phi^{(m)}_{\sigma}(t) \right|~\mathrm{d}t && \text{(\refthm{thm:nystrom-pp} w.p. $\geq 1 - \delta$)} \notag \\
        &\leq (1 + \widetilde{\varepsilon}) \int_{-1}^{1} \left| \phi_{\sigma}(t) - \phi_{\sigma}^{(m)}(t) \right|~\mathrm{d}t + \widetilde{\varepsilon} \int_{-1}^{1} \left| \phi_{\sigma}(t) \right|~\mathrm{d}t && \text{(triangle inequality)} \notag \\
        &\leq (1 + \widetilde{\varepsilon}) 2 \underline{E}_{\sigma, m} + \widetilde{\varepsilon}. && \text{(\refthm{thm:chebyshev-error}, normalization)} \notag \\
    \end{align}
    From \reflem{lem:non-negative-chebyshev-error} we see that if $0 < \sigma \ll 1$, choosing $m = \mathcal{O}(\sigma^{-1}(\log(\varepsilon^{-1}) + \log(\sigma^{-1})))$ ensures $\underline{E}_{\sigma, m} \leq \widetilde{\varepsilon}$, and thus, since $\widetilde{\varepsilon} \leq 1$,
    \begin{equation}
        \int_{-1}^{1} \left| \phi_{\sigma}(t) - \underline{\phi}_{\sigma}^{(m)}(t) \right|~\mathrm{d}t \leq (1+\widetilde{\varepsilon}) 2 \widetilde{\varepsilon} + \widetilde{\varepsilon} \leq 5 \widetilde{\varepsilon}
    \end{equation}
    from which the result follows by identifying $\varepsilon = 5 \widetilde{\varepsilon}$.
\end{proof}
